
---
title: "Penguin Body Mass Analysis"
author: "Duzhi Wu"
date: "2026-09-23"
format: html
---

## Introduction

In this post, we explore how body mass differs across three penguin species: Adelie, Chinstrap, and Gentoo.

The dataset comes from the [Palmer Penguins project](https://allisonhorst.github.io/palmerpenguins/)

The Palmer Penguins dataset is distributed under the CC0 license.

## Chunk 1: Load Data

First, we load the R packages required for our analysis.

```{r}
#| message: false
#| warning: false

library(palmerpenguins)
library(dplyr)
library(ggplot2)

data("penguins")

head(penguins)
```

The dataset contains measurements such as body mass, bill length, flipper length, species, and island.

## Chunk 2: Clean data

Before calculating summary statistics, we remove observations with missing species or body mass values.

```{r}
penguins_clean <- penguins |>
  filter(
    !is.na(species),
    !is.na(body_mass_g)
  )

cat("Original observations:", nrow(penguins), "\n")
cat("Clean observations:", nrow(penguins_clean), "\n")
```

Removing missing values ensures that our body mass calculations use observations with valid measurements.

## Chunk 3: Calculate statistics

Next, we group the observations by species and calculate the number of penguins, mean body mass, and standard deviation of body mass.

```{r}
penguin_summary <- penguins_clean |>
  group_by(species) |>
  summarise(
    count = n(),
    mean_body_mass = mean(body_mass_g),
    sd_body_mass = sd(body_mass_g),
    .groups = "drop"
  )

penguin_summary
```

These summary statistics allow us to compare the typical body mass and variability across the three species.

## Chunk 4: Visualization

We use a boxplot to compare the distribution of body mass across penguin species.

```{r}
#| label: fig-penguin-mass
#| fig-cap: "Distribution of body mass across three penguin species."
#| fig-width: 8
#| fig-height: 5

ggplot(
  penguins_clean,
  aes(x = species, y = body_mass_g, fill = species)
) +
  geom_boxplot() +
  labs(
    title = "Penguin Body Mass by Species",
    x = "Penguin Species",
    y = "Body Mass (g)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

The boxplot shows differences in the body mass distributions across species.

In conclusion, our analysis shows that Gentoo penguins have a higher average body mass than Adelie and Chinstrap penguins.
